---
title: "P8105 Homework 2"
author: "Emma Warshaw, UNI: ew2718"
date: "2022-10-01"
output: github_document
---
```{r}
library(tidyverse)
```

#Problem 1
```{r}
nyc_transit_raw = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
           col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%  
janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
  mutate(
    entry = ifelse(entry == "Yes", TRUE, FALSE) 
  )
```

Is this data tidy? 
It's not tidy.  The route numbers should have some kind of numeric value instead of the "NA" that's in there now.  It would be cleaner to have "route" as a variable and fill in information with the route number.  Need to rotate the columns from wide to long.

Answer the following questions using these data: 

How many distinct stations are there? 

Isolate for distinct pairs based on line and staion name variables: 

```{r}
nyc_transit_raw %>% 
  select(line, station_name) %>% 
  distinct
```
There are 465 distinct stations. 

How many stations are ADA compliant? 

Isolate in a similar way now, this time filtering for ada compliance.

```{r}
nyc_transit_raw %>% 
  filter(ada == TRUE) %>% 
  select(line, station_name) %>% 
  distinct
```
There are 84 stations that meet this criteria

What proportion of station entrances / exits without vending allow entrance? 

```{r}
nyc_transit_raw %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?
```{r}
nyc_transit_raw %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

nyc_transit_raw %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

#Problem 2

We'll be using an excel file, so load in the appropriate library: 

```{r}
library(readxl)
```

Read and clean the Mr. Trash Wheel sheet: 
--> (1) specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel

--> (2) use reasonable variable names

--> (3) omit rows that do not include dumpster-specific data

--> (4) round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

Use a similar process to import, clean, and organize the data for Professor Trash Wheel, and combine this with the Mr. Trash Wheel dataset to produce a single tidy dataset. To keep track of which Trash Wheel is which, you may need to add an additional variable to both datasets before combining.

Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in the resulting dataset, and give examples of key variables. For available data, what was the total weight of trash collected by Professor Trash Wheel? What was the total number of sports balls collected by Mr. Trash Wheel in 2020?
```{r}
mr_trash = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel") %>%
  janitor::clean_names() %>%
  select(-(x15:x17)) %>%
  drop_na(dumpster) %>%
  drop_na(month) %>% 
  mutate(
    sports_balls = round(sports_balls, digits = 0)) %>% 
  mutate(which_trash = "Mr. Trash") %>% 
  select(which_trash, everything())

#start on professor trash
prof_trash = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Professor Trash Wheel") %>% 
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(which_trash = "Prof. Trash") %>% 
  select(which_trash, everything())

#let's combine the data sets
trash_data = rbind(mr_trash, prof_trash)

#Calculate total weight of trash collected by Professor Trash Wheel
sum(prof_trash$weight_tons)

#Calculate the total number of sports ball collected by Mr. Trash Wheel in 2020
sum(mr_trash[which(mr_trash$year == 2020), "sports_balls"])
```

Paragraph about this data: 

Professor Trash Wheel collected a total weight of 135.5 tons of trash.  In the year 2020, Mr. Trash Wheel collected a total of 856 sports balls.  The resulting combined data frame had a total is 524 observations with 15 variables.  When cleaning this data, I decided to leave the year. month, and date columns.  Although there is redundant information in this, I think it can be useful.  For instance, trying to sum the number of sports balls collected in 2020 was made easier by still having the `year` as a variable. I could technically split the `date` variable into corresponding pieces, but for purposes of this cleaning, I didn't see the merit in additional code.  If I was doing more with this data set, I might consider it if it became functionally necessary.  

#Problem 3

First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

```{r}
pols_month = read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(col = mon, into = c('year', 'month', 'day'), sep = '-') %>%
  mutate(month = as.integer(month)) %>%
  mutate(month = case_when(
    month == 1 ~ "January",
    month == 2 ~ "February",
    month == 3 ~ "March",
    month == 4 ~ "April",
    month == 5 ~ "May",
    month == 6 ~ "June",
    month == 7 ~ "July",
    month == 8 ~ "August",
    month == 9 ~ "September",
    month == 10 ~ "October",
    month == 11 ~ "November",
    month == 12 ~ "December"
  )) %>% 
  mutate(president = prez_gop + prez_dem) %>%
  select(-prez_dem, -prez_gop, -day)
                                ### also check why certain presidents are 2
```

Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r}
snp = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(col = date, into = c('month', 'day', 'year'), sep = '/') %>% 
  mutate(month = case_when(
    month == 1 ~ "January",
    month == 2 ~ "February",
    month == 3 ~ "March",
    month == 4 ~ "April",
    month == 5 ~ "May",
    month == 6 ~ "June",
    month == 7 ~ "July",
    month == 8 ~ "August",
    month == 9 ~ "September",
    month == 10 ~ "October",
    month == 11 ~ "November",
    month == 12 ~ "December")
  ) %>% 
  select(-day) %>% 
  select(year, month, everything()) %>% 
  mutate(year_4dig = 
          if_else(year >= 50 & year <= 99, +1900), 
         if_else(year >= 00 & year <= 15), +2000) 
                                        ####the year is in two digits, have to make it 4 somehow
```

Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.
```{r}
unemployment_data = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>%
  pivot_longer(jan:dec, 
               names_to = "month",
               values_to = "unemployment_percent") %>% 
  mutate(month = case_when(
    month == "jan" ~ "January",
    month == "feb" ~ "February",
    month == "mar" ~ "March",
    month == "apr" ~ "April",
    month == "may" ~ "May",
    month == "jun" ~ "June",
    month == "jul" ~ "July",
    month == "aug" ~ "August",
    month == "sep" ~ "September",
    month == "oct" ~ "October",
    month == "nov" ~ "November",
    month == "dec" ~ "December"
  ))
```

