---
title: "p8105_hw2_ew2718"
author: "Emma Warshaw"
date: "2022-10-01"
output: github_document
---
```{r}
library(tidyverse)
```

#Problem 1
```{r}
nyc_transit_raw = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
           col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%  
janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
  mutate(
    entry = ifelse(entry == "Yes", TRUE, FALSE) 
  ) %>% 
nyc_transit_raw
```

Is this data tidy? 
It's not tidy.  The route numbers should have some kind of numeric value instead of the "NA" that's in there now.  It would be cleaner to have "route" as a variable and fill in information with the route number.  Need to rotate the columns from wide to long.

Answer the following questions using these data: 

How many distinct stations are there? 

Isolate for distinct pairs based on line and staion name variables: 

```{r}
nyc_transit_raw %>% 
  select(line, station_name) %>% 
  distinct
```
There are 465 distinct stations. 

How many stations are ADA compliant? 

Isolate in a similar way now, this time filtering for ada compliance.

```{r}
nyc_transit_raw %>% 
  filter(ada == TRUE) %>% 
  select(line, station_name) %>% 
  distinct
```
There are 84 stations that meet this criteria

What proportion of station entrances / exits without vending allow entrance? 

```{r}
nyc_transit_raw %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?
```{r}
nyc_transit_raw %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

nyc_transit_raw %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```



