---
title: "p8105_hw2_ew2718"
author: "Emma Warshaw"
date: "2022-10-01"
output: github_document
---
```{r}
library(tidyverse)
```

#Problem 1
```{r}
nyc_transit_raw = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
           col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%  
janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
  mutate(
    entry = ifelse(entry == "Yes", TRUE, FALSE) 
  ) %>% 
nyc_transit_raw
```

Is this data tidy? 
It's not tidy.  The route numbers should have some kind of numeric value instead of the "NA" that's in there now.  It would be cleaner to have "route" as a variable and fill in information with the route number.  Need to rotate the columns from wide to long.

Answer the following questions using these data: 

How many distinct stations are there? 

Isolate for distinct pairs based on line and staion name variables: 

```{r}
nyc_transit_raw %>% 
  select(line, station_name) %>% 
  distinct
```
There are 465 distinct stations. 

How many stations are ADA compliant? 

Isolate in a similar way now, this time filtering for ada compliance.

```{r}
nyc_transit_raw %>% 
  filter(ada == TRUE) %>% 
  select(line, station_name) %>% 
  distinct
```
There are 84 stations that meet this criteria

What proportion of station entrances / exits without vending allow entrance? 

```{r}
nyc_transit_raw %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?
```{r}
nyc_transit_raw %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

nyc_transit_raw %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

#Problem 2
We'll be using an excel file, so load in the appropriate library: 

```{r}
library(readxl)
```

Read and clean the Mr. Trash Wheel sheet: 
--> (1) specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel

--> (2) use reasonable variable names

--> (3) omit rows that do not include dumpster-specific data

--> (4) round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

Use a similar process to import, clean, and organize the data for Professor Trash Wheel, and combine this with the Mr. Trash Wheel dataset to produce a single tidy dataset. To keep track of which Trash Wheel is which, you may need to add an additional variable to both datasets before combining.

```{r}
mr_trash = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel") %>%
  janitor::clean_names() %>%
  select(-(x15:x17)) %>%
  drop_na(dumpster) %>%
  drop_na(month) %>% 
  mutate(
    sports_balls = round(sports_balls, digits = 0)) %>% 
  mutate(which_trash = "Mr. Trash") %>% 
  select(which_trash, everything())

#start on professor trash
prof_trash = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Professor Trash Wheel") %>% 
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(which_trash = "Prof. Trash") %>% 
  select(which_trash, everything())

#let's combine the data sets
trash_data = full_join(mr_trash, prof_trash, by = "which_trash")
```

  